---
output: github_document
---

# tidysynth

## Overview

`tidysynth` offers a tidy implementation the synthetic control method (see Abadie et al. 2003, 2010, 2015). Building on [prior packages](https://cran.r-project.org/web/packages/Synth/Synth.pdf), `tidysynth` makes a number of needed improvements when implementing the method in `R`. Increasing the user's capacity to inspect, visualize, and tune the synthetic control. 

The packages makes the following improvements:

- generates placebo synthetic controls on the fly allowing one to generate inferential statistics and plots without needing to write further code;
- offers plotting methods to easily explore the fit of the synthetic control and weights;
- relies on a nested tidy data structure with `grab_` prefix function to easily extract different aspects of the pipeline used to generate the synthetic control, increasing usability.

## Installation

```{r,eval=F}
# Install the developer version from Github
# install.packages("devtools")
devtools::install_github("tidyverse/ggplot2")
```

## Usage 

The package uses pipeline of functions to generate the synthetic control. 

| Function  | Description
|:---------:|:----------|
| `synthetic_control()` | Initialize a `synth_tbl`. The main initialization function of the `tidysynth` method. |
| `generate_predictor()` | Create one or more scalar variables summarizing covariate data across a specified time window. These predictor variables are used to fit the synthetic control. |
| `generate_weigths()` | fits the unit and predictor weights used to generate the synthetic control. |
| `generate_control()` | generates the synthetic control using the optimized weights. |

The following example comes from Abadie et. al 2010, which evaluates the impact of Proposition 99 on cigarette consumption in California. Proposition 99 greatly increased funding to anti-smoking campaigns and imposed a 25 cent tax on a pack og cigarettes
```{r,message=F}
require(tidysynth)
data("smoking")
smoking %>% dplyr::glimpse()
```

Let's generate a synthetic California using information from the "donor pool", which is composed of control states where a law like Proposition 99 wasn't implemented. This is the pool of potential cases from which we borrow information when fitting to the observed California in the pre-intervention period.

```{r,cache=T}
smoking_out <-
  
  smoking %>%
  
  # initial the synthetic control object
  synthetic_control(outcome = cigsale, # what variable are we generating the synthetic control for?
                    unit = state, # Unit index in the panel data
                    time = year, # Time index in the panel data
                    i_unit = "California", # Name of the unit where the intervention occurred
                    i_time = 1988, # time period when the intervention occurred
                    generate_placebos=T # generate placebo synthetic controls as well
                    ) %>%
  
  # Generate the aggregate predictors used to fit the weights
  generate_predictor(time_window=1980:1988,
                     ln_income = mean(lnincome, na.rm = T),
                     ret_price = mean(retprice, na.rm = T),
                     youth = mean(age15to24, na.rm = T)) %>%
  
  # What was the averate beer consumption in the donor pool from 1984 - 1988
  generate_predictor(time_window=1984:1988,
                     beer_sales = mean(beer, na.rm = T)) %>%
  
  # Lagged cigarette sales
  generate_predictor(time_window=1975,
                     cigsale_1975 = cigsale) %>%
  generate_predictor(time_window=1980,
                     cigsale_1980 = cigsale) %>%
  generate_predictor(time_window=1988,
                     cigsale_1988 = cigsale) %>%
  
  
  # Generate the fitted weights for the synthetic control
  generate_weights(optimization_window =1970:1988, # Can set the time window in the 
                   Margin.ipop=.02,Sigf.ipop=7,Bound.ipop=6) %>%
  
  # Generate the synthetic control
  generate_control()
```

Once the synthetic control is generated, one can easily assess the fit... 

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_trends()
```


...and the post-intervention period difference, which reflects the causal effect (i.e. the difference between the observed and counterfactual).

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_differences()
```

In addition, one can easily examine the weighting of the units and variables in the fit. This allows one to see which cases were used, in part, to generate the synthetic control

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_weights()
```

Another useful way of evaluating the synthetic control is to look at how comparable the control is to the covariates of the actual fit.

```{r}
smoking_out %>% grab_balance_table()
```


### Inference 

The method relies on generating placebo synthetic controls for every donor in the donor pool exactly as was done for California. By setting `generate_placebos = TRUE` when initializing the `synth_tbl` with `synthetic_control()`, placebo cases are generated alongside the target unit, where the intervention actually took place. The aim is then to compare the behavior between the target difference and the placebo differences.

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_placebos()
```

Note that the `plot_placebos()` function automatically prunes placebo synthetic controls that poorly fit and throw off the scale when plotting the placebos. To prune, the function looks at the pre-intervention period mean squared prediction error (MSPE) (i.e. a metric that reflects how well the synthetic control maps to the observed outcome time series). If a placebo control has a MSPE that is two times beyond the target case, then it's dropped. To turn off this behavior, set `prune = FALSE`. 

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_placebos(prune=F)
```

Finally, Adabie et al. 2010 outline a way of constructing Fisher's Exact P-value but taking dividing the post-intervention period MSPE by the pre-intervention period MSPE and then ranking all the cases by this ratio. The p-value is the constructed by taking the rank/total.^[Note this implies that you'd need at least 20 cases in the donor pool to get a conventional p-value (.05).] The idea is that if the synthetic control fits the observed time series well (low MSPE in the pre-period) and diverges in the post-period (high MSPE in the post-period) then there is a meaningful effect due to the intervention. Put differently, a good fit in the pre-period means we've generate a descent mapping of the cases behavior. If the intervention had no effect, then the post-period and pre-period should continue to map onto one another fairly well. 

This ratio can be easily plotted, offering insight into the rarity of the case where the intervention actually occurred. Here one can see that California is a clear outlier.

```{r,fig.align="center",fig.width=10,fig.height=5,dpi=300}
smoking_out %>% plot_mspe_ratio()
```

For most specific information, there is a significance table that can be extracted with one of the many `grab_` prefix functions.
```{r}
smoking_out %>% grab_signficance()
```


### Accessing Data 

In addition, to the main data pipeline for generating the synthetic control and the `plot_` prefix functions for visualizing the output, there are also a number of `grab_` prefix functions that allow one easily to grab the data contained within a `synth_tbl` object.

Note that at its core, a `synth_tbl` is just a tidy nested data frame, where each component of the synthetic control pipeline is appended on as the function is fun. 

```{r}
smoking_out
```

To make it easier to access the relevant data fields (as writing out the relevant unnesting procedure can get involved), the `grab_` prefix functions come into play. 

| Function | Description |
|:------------------:|:--------------------|
| `grab_outcome()` |  Extract the outcome variable generated by `synthetic_control()`. |
| `grab_predictors()` |  Extract the aggregate-level covariates generated by `generate_predictor()`. |
| `grab_unit_weights()` |Extract the unit weights generated by `generate_weights()`. |
| `grab_predictor_weights()` | Extract the predictor variable weights generated by `generate_weights()`. | 
| `grab_loss()` | Extract the RMSE loss of the optimized weights generated by `generate_weights()`. | 
| `grab_synthetic_control()` | Extract the synthetic control generated using `generate_control()`. |
| `grab_signficance()` | Generate inferential statistics comparing the rarity of the unit that actually received the intervention to the placebo units in the donor pool. |
| `grab_balance_table()` | Compare the distributions of the aggregate-level predictors for the observed intervention unit, the synthetic control, and the donor pool average.  | 


```{r}
smoking_out %>% grab_synthetic_control()
```

Note that most all the `grab_` functions allow for extraction of the placebo units as well. 
```{r}
smoking_out %>% grab_synthetic_control(placebo=T)
```
